---
title: "Lecture 12"
author: "Grace Lee"
date: "May 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 1.1

load bio3D package and download 1HSG pdb file
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

Read pdb structure
```{r}
hiv <- read.pdb(file.name)
```

##Q1
non-protein residue: water and MK1
What does resid correspond to and how would you get a listing of all reside values in this structure?
resid -> residuals
```{r}
hiv$atom$resno
```

# Section 1.2

Select out protein and ligand only
```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")
prot
lig
```

Write out these new pdb files
```{r}
write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, "1hsg_ligand.pdb")
```

# Section 1.3
done on class computer (b/c could not downlad ADT from MGLTools on PC)

## Q3
charges make sense

# Section 1.4
done on class computer

# Section 1.5
created config.txt file

# Section 2.3
processing docking results for VMD display
```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
write.pdb(res, "results.pdb")
```

Select non-hydrogen atoms from our docking result file
```{r}
inds.res<- atom.select(res, "noh")
inds.res
inds.MK <- atom.select(lig, "noh")
inds.MK
```


```{r}
#res$xyz[, inds.res$xyz]
rmsd(lig, res$xyz[ ,inds.res$xyz])
```

