---
title: "Lecture 14"
author: "Grace Lee"
date: "May 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Trascriptomics and the analysis of RNA-seq data

Import countData and colData into R
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```

# Toy differential gene expession
```{r}
View(metadata)
```

```{r}
control <- metadata[metadata[,"dex"] == "control",]
control.mean <- rowSums( counts[ ,control$id])/length(control$id)
names(control.mean) <- counts$ensgene
```

**Q1**
divided by 4 always even when adding more samples
```{r}
table(metadata$dex)
length(control$id)
```

**Q2**
```{r}
treated <- metadata[metadata[, "dex"] == "treated",]
treated.mean <- rowSums( counts[ ,treated$id])/length(treated$id)
names(treated.mean) <- counts$ensgene
```

```{r}
meancounts <- data.frame(control.mean, treated.mean)
colSums(meancounts)
```

**Q3**
```{r}
plot(control.mean, treated.mean, xlab = "Control", ylab = "Treated")
# only see some points because the points overlap
```


```{r}
plot(control.mean, treated.mean, log = "xy", xlab = "log Control", ylab = "log Treated")
```

Calculate log2foldchange, add it to out meancounts data.frame
```{r}
meancounts$log2fc <- log2(meancounts[, "treated.mean"]/meancounts[, "control.mean"])
head(meancounts)
```

```{r}
zero.vals <- which(meancounts[, 1:2]==0, arr.ind= TRUE)
# arr.ind --> (array.index) give row and column index
# inspect below for example
  #x <- matrix(c(1:10), ncol = 2, byrow = 1)
  #x[5,2] <- 0
  #x
  #which(x ==0)
  #which(x ==0, arr.ind = TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
                
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
head(up.ind)
head(down.ind)
```

**Q5** 
```{r}
sum(up.ind == "TRUE")
sum(down.ind =="TRUE")
```

```{r}
paste("Number of UP genes:", sum(up.ind))
paste("Number of DOWN genes:", sum(down.ind))
```

# Adding annotation data
```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

**Q6**
```{r}
merge(mycounts, anno, by.x = "row.names", by.y = "ensgene")
```


```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```

**Q7**
```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")
```

**Q8**

Load Bioconductor's annotation packages
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
head(mycounts[up.ind,])
```

# DeSeq2 analysis
```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

Running DESeq pipelin on the dds object
```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
summary(res)
```

```{r}
resOrdered <- res[order(res$pvalue),]
```

```{r}
res05 <- results(dds, alpha = 0.05)
summary(res05)
```

**Q9**
How many are significant with an adjusted p-value < 0.05?
```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)
```
 
 **Q9**
How many are significant with an adjusted p-value < 0.01?
```{r}
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

**Q10**
Add annotation from previous anno object using merge() function
```{r}
merge(resSig01, anno, by.x = "row.names", by.y = "ensgene")
```

Add annotation to resSig01 using mapIds() function
```{r}
head(resSig01)
```

```{r}
resSig01$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
resSig01$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
```
```{r}
resSig01$uniprot<- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
resSig01
```

```{r}
ord <- order( resSig01$padj)
head(resSig01[ord,])
```

Write out ordered significant results w/ annotations
```{r}
write.csv(resSig01[ord,], "signif01_results,csv")
```

# Data Visualization
```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]
```
```{r}
rownames( resSig01[i,])
```

```{r}
plotCounts(dds, gene = "ENSG00000103196", intgroup = "dex")
```

```{r}
d <- plotCounts(dds, gene="ENSG00000103196", intgroup = "dex", returnData = TRUE)
head(d)
```

```{r}
boxplot(count ~dex, data = d)
```

```{r}
library(ggplot2)
ggplot(d, aes(dex, count)) + geom_boxplot(aes(fill = dex)) + scale_y_log10() +ggtitle("CRISPLD2")
```

#MA & Volcano plots
```{r}
res$sig <- res$padj < 0.05
table(res$sig)
```

```{r}
sum(is.na(res$sig))
```

```{r}
plotMA(res, ylim=c(-2,2))
```

```{r}
resLFC <- lfcShrink(dds, coef=2)
resLFC
```

# Volcano plor
```{r}
ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```

