---
title: "Lecture 16"
author: "Grace Lee"
date: "May 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(readr)
```

Read data
```{r}
exp_data <- read.csv("Experimental_results.csv")
head(exp_data)
```

Remove column "X" from a dataframe
can also do [,-1]
select() not from base R, part of the dplyr package
```{r}
exp_data <- select(exp_data, -X)
head(exp_data)
```

Select function practice
```{r}
Mut_str_env <- select(exp_data, Mut_ID, Strain, Environment)
head(Mut_str_env)
```

Rearrange data
want to plot time vs frequency using ggplot
time and frequency are in rows in the initial data
want to create time and frequency in columns
```{r}
exp_rearranged <- gather(exp_data, Generation, Frequency, H_0:H_96)
head(exp_rearranged)
```

Separate values in "Generation" column into 2 columns
```{r}
table_for_graph <- separate(exp_rearranged, Generation, into=c("H", "Time"))
head(table_for_graph)
```

Remove columns H
```{r}
table_for_graph <- select(table_for_graph, -H)
head(table_for_graph)
```

Remove NA values
```{r}
table_cleaned <- na.omit(table_for_graph)
table_cleaned$Time <- as.numeric(table_cleaned$Time)
head(table_cleaned)
```

Plot data for each strain separately
```{r}
DivAnc <- filter(table_cleaned, table_cleaned$Strain == "DivAnc")
L013 <- filter(table_cleaned, table_cleaned$Strain == "L013")
```

Make a plot for DivAnc strain
```{r}
DivAnc_plot = ggplot(DivAnc) + geom_line(aes(x = Time, y= Frequency, group = BC_ID), alpha = 0.2, colour = "#000033") +ggtitle("DivAnc_SC3") + theme(plot.title = element_text(hjust = 0.5)) +xlab("Time, hours") + ylab("Log10(Barcode frequency)")
DivAnc_plot
```


Make a plot for L013 Strain
```{r}
L013_plot = ggplot(L013) + geom_line(aes(x = Time, y= Frequency, group = BC_ID), alpha = 0.2, colour = "#CC6633") +ggtitle("L013_SC3") + theme(plot.title = element_text(hjust = 0.5)) +xlab("Time, hours") + ylab("Log10(Barcode frequency)")
L013_plot
```

Make 2 plots at the same time
```{r}
ggplot(table_cleaned) + geom_line(aes(x= Time, y= Frequency, group = BC_ID), alpha = 0.2, colour = "#000033") +facet_grid(.~Strain) +ggtitle("Barcode trajectories") +theme(plot.title = element_text(hjust = 0.5)) +xlab("Times, hours") + ylab("Log10(Barcode frequency)")
```

Pick one mutation and check how it behaves in different strains
pick Mut_ID == 34
```{r}
mut34 <- filter(table_cleaned, table_cleaned$Mut_ID == "34")
mut34
```

```{r}
ggplot(mut34,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

Filter out barcodes with frequency > (-5)
because seqeuncing frequency barcode has low coverage
```{r}
mut34_f <- filter(mut34, mut34$Frequency> (-5))
mut34_f
```

Plot filtered data
```{r}
ggplot(mut34_f,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```

Same plot different methods
```{r}
ggplot(mut34_f,aes(Time, Frequency, colour = BC_ID, group=BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+facet_grid(.~Strain)+theme(legend.position="none")+ggtitle(paste("Mutation",34, sep="_"))+xlab("Time, hours")+ ylab("Log10(Barcode frequency)")
```

Estimate slope for each barcode
picked BC_25361
```{r}
BC_25361<-filter(mut34_f, mut34_f$BC_ID=="25361", mut34_f$Strain=="DivAnc")
BC_25361
```

Plot frequency of this barcode
```{r}
BC_plot = ggplot(BC_25361, aes(Time, Frequency, colour = BC_ID))+geom_point() +theme(legend.position = "none") +ggtitle("BC_25361")+xlab("Time, hours") +ylab("Log10(Frequency)")
BC_plot
```

Use lm function (linear model function) to fit the line to these points
```{r}
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```
#shortcut
```{r}
BC_plot_lm = BC_plot + geom_smooth(se = FALSE, method = "lm")
BC_plot_lm
```

What does lm function return
```{r}
regression_model <- lm(Frequency~Time, BC_25361)
summary_data <- summary(regression_model)
summary_data
```

Interested in Slope and Intercept
```{r}
#Time Slope
Time <- summary_data$coefficients[2]
Time
#Intercept
Intercept <- summary_data$coefficients[1]
Intercept
```

Find slopes for each barcode
```{r}
data_header = matrix(data = NA, nrow = 1, ncol = 7)
        data_header[1]="Mut_ID"
        data_header[2]="BC_ID"
        data_header[3]="Strain"
        data_header[4]="Slope"
        data_header[5]="Intercept"
        data_header[6]="R^2"
write.table(data_header, "~/Documents/UCSD/2018 Spring Quarter/BIMM 143/BIMM143_github/Lecture 16/Tnseq_practice_output.csv", append = FALSE, sep = ",", eol = "\n", dec = ".", row.names = FALSE, col.names = FALSE)
```

