---
title: "9-Exercise (PCA)"
author: "Grace Lee"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
wisc.df
```

#Convert the features of the data: wisc.data
```{r}
wisc.data <- as.matrix(wisc.df)
wisc.data
```

#Set the row names of wisc.data
```{r}
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

#Create diagnosis vector by completing the missing code
```{r}
table(wisc.df$diagnosis)
```
```{r}
wisc.df$diagnosis == "M"
# R set TRUE = 1
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
diagnosis
table(diagnosis)
```

# How many observations are in this data set
```{r}
nrow(wisc.df)
```

# How many variables/features in the data are suffixes with_mean?
```{r}
# Return columns with word 'mean' in them
grep("mean", colnames(wisc.data), value = TRUE)

# Return columns without word 'mean'
grep("mean", colnames(wisc.data), value = TRUE, invert = TRUE)#gives opposite of above code

# Return the position (ie index) of things with 'mean' in them
grep("mean", colnames(wisc.data))

# Return total number of columns with 'mean' in them
length(grep("mean", colnames(wisc.data)))
```

# How many of the observations have a malignant diagnosis?
```{r}
table(diagnosis)
sum(diagnosis) # can do sum here because previous assigned "M" = 1, and "B" = 0
```

#Convert the features of the data: wisc.data
```{r}
wisc.data <- as.matrix(wisc.df[,3:(ncol(wisc.df)-1)])
# did (ncol(wisc.df)-1) *(minus 1) because later when performing PCA, NA in the X column is not allowed
  # or we can do
  # wisc.data <- as.matrix(wisc.df[,3:(ncol(wisc.df)])
  # wisc_data <- wisc.data[,-31] 
  # gives every column but the 31st column (column X)
  # 31, b/c when we did as.matrix() we extracted 3rd to 33rd column which became 1st to        31st column in the new matrix
rownames(wisc.data) <- wisc.df$id
head(wisc.data)
```

# Check if need to scale PCA data
```{r}
colMeans(wisc.data)
apply(wisc.data, 2, sd)
```

# Check data
```{r}
head(wisc.data)
```

#Perform PCA on wisc.data
```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
```

# Look at summary of results
```{r}
summary(wisc.pr)
```

# What is in your PCA output
```{r}
attributes(wisc.pr)
```

#Q4 From your results, what proportion of the original variance is captured by the first principal components (PC1)?
# ANS: 0.4427

# Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
# ANS: 3

#Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
# ANS: 7

# Create biplot
```{r}
biplot(wisc.pr)
```

# Look at PC values
```{r}
wisc.pr$x
```

# Scatter plot observations by cmponents 1 and 2
# color by diagnosis vector (malignant vs benign)
```{r}
plot(wisc.pr$x[, c(1, 2)], col = (diagnosis +1), pch = as.vector(wisc.df$diagnosis),xlab = "PC1", ylab = "PC2")
```

# Scatter plot observations by components 1 and 3
# did diagnosis + 1 because 0 is white 
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[, 3], col = (diagnosis +1), xlab = "PC1", ylab = "PC3")
```
# between plots of pc1 vs pc2 and pc1 vs pc3, we see that the first plot (pc1 vs pc2) provides a clearer cut between the two clusters. This is because pc2 provides a larger variance

# Calculate percent variance
```{r}
wisc.var <- wisc.pr$sdev^2
pve <- (wisc.var)/sum(wisc.var)
pve
```

# Alternatice scree plot of the same data, note data driven y-axis
```{r}
barplot (pve, ylab = "Percent of Variance Explained", names.arg=paste0("pc", 1:length(pve)), las=2, axes = FALSE, main = "Scree Plot")
# axis is turned off
# so we create axis using the code below
axis(2, at=pve, labels=round(pve,2)*100)
```

# Plot variance explained for each principal component
```{r}
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = "o")
```

# Plot cumulative proportion of variance for each princpal component
```{r}
plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = "o")
```

```{r}
par(mfrow = c(1,2))
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = "o")
plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = "o")
```

