---
title: "Lecture 13"
author: "Grace Lee"
date: "May 15, 2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 1: 1000 Genomes Data Analysis
Downloaded genotype data from Ensemble for Asthma associated SNP rs8067378 
See: (MXL) Mecican Ancestry in Los Angeles, California

## Question 5
Read CSV file
```{r}
genotype <- read.csv("Lecture 13 data 373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(genotype)
```

Find how many of each genotype there are in this dataset
```{r}
genotype.summary <- table(genotype[,2])
genotype.summary
```

Proportion of each genotype
```{r}
genotype.summary/nrow(genotype) * 100
```

**Q5** What proportion of the Mexican Ancestry in Los Angeles sample population (MXL)
are homozygous for the asthma associated SNP (G|G)?
Ans: 14%

# Section 2: RNA - Seq
Look more closely at sample HG00109 with G|G genotype for this SNP (from GBR population)
Use http://149.165.168.199/

From Lecture 13 Slide #58
Side-note on Fastq quality scores
```{r}
#install.packages("seqinr")
#install.packages("gtools")
library(seqinr)
library(gtools)
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@") ) - 33
phred
```

```{r}
prob <- 10**(-phred/10)
prob
```

# Section 4
```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

**Q14** Sample size for each genotype
```{r}
summary(expr)
table(expr$geno)
```

How many samples?
```{r}
nrow(expr)
```

```{r}
summary( expr$exp[ (expr$geno == "G/G")])
summary( expr$exp[ (expr$geno == "A/A")])
summary( expr$exp[ (expr$geno == "A/G")])
```

Boxplot (genotype vs expression levelss)
```{r}
boxplot(exp ~ geno, data = expr, notch = TRUE)
```

Use ggplot
```{r}
library(ggplot2)
```

```{r}
ggplot(expr, aes(geno, exp)) + geom_boxplot()
```

Histogram of the exp column with ggplot2
```{r}
ggplot(expr, aes(exp, fill = geno)) + geom_density()
```

```{r}
# Boxplot with the data shown
ggplot(expr, aes(geno, exp, fill=geno)) + 
  geom_boxplot(notch=TRUE, outlier.shape = NA) + 
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4)
```

