---
title: "Lecture 11 - new"
author: "Grace Lee"
date: "May 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1
## PDB statistics

# Q1
Download CSV file from PDB database. Read this into R and determine fraction of X-ray structures.
```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Lets calculate something
```{r}
percent <- (pdb.stats$Total/ sum(pdb.stats$Total))*100
names(percent) <- pdb.stats$Experimental.Method
percent
```

# Q2 Ans: 118 (found on webpage)

## Section 3
# Working with Bio3D

# Section 3.1
Load bio3D
```{r}
library(bio3d)
```

List function of bio3D package

# Section 3.2
To see example of functions
```{r}
#example("plot.bio3d")
```

# Section 3.3
Read 1hsg
```{r}
pdb <- read.pdb("1hsg")
pdb
```
# Question 6
172 non-protein residues (HOH and MK1)
1514 protein residues

# Question 6
```{r}
sum(pdb$atom$resid == "MK1") + sum(pdb$atom$resid == "HOH")
nrow (pdb$atom) - (sum(pdb$atom$resid == "MK1") + sum(pdb$atom$resid == "HOH"))
```
non-protein residues: water & MK1
aa residues: 1514

# Section 3.4
```{r}
attributes(pdb)
head(pdb$atom)

# Print a subset of $atom data fro the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x", "y","z")]

# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]

# Which allows us to do the following
plot.bio3d(pdb$atom$b[pdb$calpha], sse = pdb, typ = "l", ylab = "B-factor")
```
# Question 7
```{r}
str(pdb$atom)
```
# Question 7
data frame


# Print a summary of the coordinate data in $xyz
```{r}
pdb$xyz
```

# Examine the row and column dimensions
```{r}
dim(pdb$xyz)
```

# Print coordinates for the first two atom
```{r}
pdb$xzy[ 1, atom2xyz(1:2ss)]
```

# Section 3.5
# Select all C-alpha atoms (return their indices)
```{r}
inds <- atom.select(pdb, "calpha")
inds
```

# Print details of the first few selected atoms
```{r}
head(pdb$atom[inds$atom,])
```

# Print details of the first few selected xyz coordinates
```{r}
head(pdb$xyz[ , inds$xyz])
```

# Select other strings
```{r}
# Select chain A
a.inds <- atom.select(pdb, chain = "A")
# Select C-alphas of chain A
ca.inds <- atom.select(pdb, "calpha", chain = "A")
# combine multiple selection criteria to return their intersection
cab.inds <- atom.select (pdb, elety= c("CA", "CB"), chain = "A", resno = 10:20)
```

# Question 8
Looks good so lets continue to trim our PDB and write a new C-alpha only PDB file for viewing in VMD
```{r}
ca.pdb <- trim.pdb(pdb, inds=inds)
ca.pdb
#create file
write.pdb(ca.pdb, file = "calpha_only.pdb")
```


# Section 4
Download some example PDB files
```{r}
ids <- c("1TND_B", "1AGR_A", "1GG2_A", "1KJY_A", "4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

Extract and align the chains we are interested in
```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile = "C:/Users/Gracie/Downloads/muscle.exe")
pdbs
```

# Question 8
fit = TRUE in the rmsd() function performs coordinate superpostition before RMSD calculation

# Section 5

#Get aa seq of 1AKE
```{r}
aa <- get.seq("1ake_A")
aa
```

# Blast or hmmer search
# cache = TRUE -> stores the result
```{r, cache= TRUE}
b <- blast.pdb(aa)
```

# Plot a summary of search resuls
```{r}
hits <- plot(b)
# a quantified view of similarity between the query and the results
```

```{r}
attributes(hits)
head(hits$pdb.id) #view of the most similar proteins
```

# Fetch files
```{r}
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
```
#Align structures
```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile = "C:/Users/Gracie/Downloads/muscle.exe")
```

# Vector containing PDB codes
```{r}
ids <- basename.pdb(pdbs$id)
```

# Draw schematic alignment
```{r}
plot(pdbs, labels = ids)
```



